---
title: "Tables and Figures for 'An empirically driven and inclusive inventory facilitates recognition
of diverse area-based conservation of nature' "
author: "Yifan (Flora) He"
date: "3/11/2024"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flextable)
library(tidyverse)
library(janitor)
library(readxl)
library(paletteer)
library(RColorBrewer)
```


#  import data
Note: all units are in square km
```{r}
ctr_sys <- read_csv("data/total_area_by_country_and_subfam.csv")
land_sys <- read_csv("data/land_area_by_subfam.csv")
land_ctr <- read_csv("data/land_conserved_by_country.csv")
biome_sys <-read_csv("data/biome_area_by_subfam.csv")
biome_ctr <-read_csv("data/biome_area_by_country.csv")
ctr_sys_land <- read_csv("data/land_area_by_country_and_subfam.csv")
ctr_sys_biome <- read_csv("data/biome_area_by_country_and_subfam.csv")

totalland <- read_csv("data/total_land_area.csv")
land_by_ctr <- read_csv("data/land_area_by_country.csv")
biome_by_ctr <- read_csv("data/biome_by_country.csv")
non_poly <- read_excel("data/Non_polygon_summary.xlsx") # didn't change this one 

total_conserved <-read_csv("data/total_conserved.csv")
land_conserved <-read_csv("data/total_conserved_land.csv")
biome_conserved <-read_csv("data/total_conserved_biome.csv")

overlap_area <- read_csv("data/overlap_matrix.csv")

wdpa2016_matrix <- read_csv("data/wdpa2016_matrix.csv")
atlas_matrix <- read_csv("data/atlas_matrix.csv")

wdpa_wdoecm_2023_matrix <- read_csv("data/wdpa_wdoecm_2023_matrix.csv")

```




# Main text Figure 5. Comparison of data covered in our inventory and data reported to WDPA in 2016

## WDPA categorization matrix 
```{r}
wdpa2016_matrix <- wdpa2016_matrix %>% 
  mutate_all(as.character)

wdpa2016_matrix <- wdpa2016_matrix %>% 
  mutate_all(~ case_when(
    .== "10" ~ "Polygon",
    .== "11"~ "Polygon and point",
    .== "1" ~ "Point",
    .default = as.character(.)))

# add empty columns of subfamilies not reported
wdpa2016_matrix$"Area under community-based natural resources management" <- NA
wdpa2016_matrix$"Area under payment for ecosystem services program" <- NA
wdpa2016_matrix$"Buffer zone" <- NA
wdpa2016_matrix$"Conservation agreement" <- NA
wdpa2016_matrix$"Conservation concession" <- NA
wdpa2016_matrix$"Eco-certified production area" <- NA
wdpa2016_matrix$"Military training ground" <- NA
wdpa2016_matrix$"Private recreational area" <- NA

colorder <- c("ISO3",
              "Area under community-based natural resources management",
              "Area under payment for ecosystem services program", 
              "Buffer zone", 
              "Conservation agreement",
              "Conservation concession", 
              "Eco-certified production area", 
              "Indigenous territory", 
              "IPLC-led conservation area", 
              "Military training ground", 
              "Private recreational area", 
              "Privately protected area",
              "State-designated protected area", 
              "State-designated recreational area",
              "State-required sustainable forest production area", 
              "Other",
              "non-BPA",
              "International designation",
              "potential BPA not included in Atlas",
              "TBD",
              "Proposed PA")

wdpa2016_matrix <- wdpa2016_matrix %>% 
  select(all_of(colorder))

wdpa2016_matrix_transpose <- t(wdpa2016_matrix) 
wdpa2016_matrix_transpose <- wdpa2016_matrix_transpose%>% 
  as.data.frame(row.names(wdpa2016_matrix_transpose))
# write.csv(wdpa2016_matrix_transpose, "data/wdpa2016_matrix_cleaned.csv")

t1 <- wdpa2016_matrix %>% 
  flextable()
t1 <- theme_vanilla(t1)
t1
```

## Our inventory's categorization matrix
```{r}
atlas_matrix <- atlas_matrix %>% 
  mutate_all(as.character)

atlas_matrix <- atlas_matrix %>% 
  mutate_all(~ case_when(
    .== "10" ~ "Polygon",
    .== "11"~ "Polygon and point",
    .== "10.1" ~ "Polygon and statistics",
    .== "1" ~ "Point",
    .=="0.1" ~ "Statistics",
    .default = as.character(.)))

atlas_matrix$"International designation" <- NA
atlas_matrix$"potential BPA not included in Atlas" <- NA
atlas_matrix$"non-BPA" <- NA
atlas_matrix$"TBD"<- NA
atlas_matrix$"Proposed PA" <- NA

colorder <- c("ISO_CODE",
              "Area under community-based natural resources management",
              "Area under payment for ecosystem services program", 
              "Buffer zone", 
              "Conservation agreement",
              "Conservation concession", 
              "Eco-certified production area", 
              "Indigenous territory", 
              "IPLC-led conservation area", 
              "Military training ground", 
              "Private recreational area", 
              "Privately protected area",
              "State-designated protected area", 
              "State-designated recreational area",
              "State-required sustainable forest production area", 
              "Other",
              "non-BPA",
              "International designation",
              "potential BPA not included in Atlas",
              "TBD",
              "Proposed PA")
atlas_matrix <- atlas_matrix %>% 
  select(all_of(colorder)) %>% 
  arrange(ISO_CODE)

# transpose
atlas_matrix_transpose <- t(atlas_matrix)
atlas_matrix_transpose <- atlas_matrix_transpose %>% 
  as.data.frame(row.names(atlas_matrix_transpose))
  

# write.csv(atlas_matrix_transpose, "data/atlas_matrix_cleaned.csv", row.names = T)

t2 <- atlas_matrix %>% 
  flextable()
t2 <- theme_vanilla(t2)
t2

# manually transpose and combine with WDPA matrix in Excel. used those instead of t1 & t2 in the manuscript. 

```


# Figure S1. Percentage of area-based conservation governance system subfamily within the total land area by country.

```{r }

land_by_ctr <- rename(land_by_ctr, land_area =  area_geo, country = ISO_A3)

head(ctr_sys_land)
names(ctr_sys_land)[2] <- "Country"
names(ctr_sys_land)[1] <- "System" 
ctr_sys_land_bar <- ctr_sys_land %>%
  clean_names() %>%
  select(country, system, area_geo) %>% 
  filter(country %in% c("BOL", "BRA", "COL", "ECU", 
                        "GUF", "GUY", "PER", "SUR", "VEN"))%>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%  # convert to 2 digits
  right_join(land_by_ctr, by = "country")

# create new column of categories containing small contributors
ctr_sys_land_bar$system_sm <- ctr_sys_land_bar$system
ctr_sys_land_bar$system_sm[ctr_sys_land_bar$system %in% 
                             c("Private recreational area", 
                               "Conservation concession", 
                               "Conservation agreement", 
                               "Privately protected area", 
                               "State-designated recreational area")] <-
  "Small contributors"


# calculate percent land
ctr_sys_land_bar$percent_land <- 
  ctr_sys_land_bar$area_geo/ctr_sys_land_bar$land_area 


# removed non BPA.
ctr_sys_land_bar <- 
  ctr_sys_land_bar[!ctr_sys_land_bar$system == "non-BPA",] 

# get percent area by system_sm: 
bar_fig <- ctr_sys_land_bar %>% group_by(system_sm, country) %>% 
  summarise(percent_land_sm = sum(percent_land))

percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}
bar_fig$percent_land_label <-
  percent(bar_fig$percent_land_sm)

#reorder sytems
bar_fig$system_sm <- as.factor(bar_fig$system_sm)
levels(bar_fig$system_sm)
bar_fig$system_sm <- factor(bar_fig$system_sm, 
                                  levels = c("Area under community-based natural resources management",
                                             "Area under payment for ecosystem services program",
                                             "Buffer zone",
                                             "Eco-certified production area",
                                             "Indigenous territory",
                                             "IPLC-led conservation area",
                                             "Military training ground",
                                             "State-designated protected area",
                                             "State-required sustainable forest production area",
                                             "Small contributors",
                                             "Other"))
levels(bar_fig$system_sm) # check output

# Assign colors
amazoncol_sm <- c("#8c6bb1", # "Area under community-based natural resources management"
               "#74a9cf", #"Area under payment for ecosystem services program"
               "#AEE171", #"Buffer zone"
               "#CD6E6E", # "Eco-certified production area"
               "#6677CD", # "Indigenous territory"
               "#6A3D9A", #"IPLC-led conservation area",
               "#CC6428", #"Military training ground"
               "#feb24c",#"small contributor"
               "#006837", #"State-designated protected area"
               "#fcc5c0", # "State-required sustainable forest production area"
               "#CCCCCC" #"Other"
)
# Amazon boundary: #686868
# boundary color: #828282 (50% gray)
# country background: white 

names(amazoncol_sm) <- c(
  "Area under community-based natural resources management", 
  "Area under payment for ecosystem services program",
  "Buffer zone",
  "Eco-certified production area",
  "Indigenous territory",
  "IPLC-led conservation area",
  "Military training ground",
  "Small contributors",
  "State-designated protected area",
  "State-required sustainable forest production area",
  "Other") 


# Turn ISO into country name

bar_fig$country <- factor(bar_fig$country, 
                          levels = c("BOL", "BRA", "COL", "ECU", 
                                     "GUF", "GUY", "PER", "SUR", "VEN"),
                          labels = c("Bolivia", "Brazil", "Colombia", 
                                     "Ecuador", "French Guiana", "Guyana", "Peru",
                                     "Suriname", "Venezuela"))

ggplot(bar_fig, aes(x = system_sm, y = percent_land_sm, fill = system_sm))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = amazoncol_sm)+
  facet_wrap(~country, nrow = 3, ncol = 3)+
  geom_text(aes(label = percent_land_label), 
            position = position_dodge(0.9),
            vjust = - 0.5,
            size = 2.5)+
  ylab("Percent of country's total land area")+
  guides(fill=guide_legend(title="System subfamily"))+ # change legend title 
  scale_y_continuous(labels = scales::percent)+
  theme(axis.title.x=element_blank(), # remove x labels - too long 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

 

```

# Figure S2. Percentage of area-based conservation governance system subfamilies within the Amazon biome area by country. 

```{r eval}
biome_by_ctr <- rename(biome_by_ctr, country = ISO_A3, biome_area =  area_geo)

head(ctr_sys_biome)
names(ctr_sys_biome)[2] <- "Country"
names(ctr_sys_biome)[1] <- "System" 
ctr_sys_biome_bar <- ctr_sys_biome %>%
  clean_names() %>%
  select(country, system, area_geo) %>% 
  filter(country %in% c("BOL", "BRA", "COL", "ECU", "GUF", "GUY", "PER", "SUR", "VEN"))%>%
  mutate(across(where(is.numeric), round, digits = 2)) %>% 
  right_join(biome_by_ctr, by = "country")

# create new column of categories containing small contributors
ctr_sys_biome_bar$system_sm <- ctr_sys_biome_bar$system
ctr_sys_biome_bar$system_sm[ctr_sys_biome_bar$system %in% 
                             c("Private recreational area",  
                               "Conservation concession", 
                               "Conservation agreement", 
                               "Privately protected area", 
                               "State-designated recreational area")] <-
  "Small contributors"

# calculate percent biome
ctr_sys_biome_bar$percent_biome <- ctr_sys_biome_bar$area_geo/ctr_sys_biome_bar$biome_area 


# removed non-BPA 
ctr_sys_biome_bar <- ctr_sys_biome_bar[!ctr_sys_biome_bar$system == "non-BPA",] 

# get percent area by system_sm: 
bar_fig_biome <- ctr_sys_biome_bar %>% group_by(system_sm, country) %>% 
  summarise(percent_biome_sm = sum(percent_biome))

percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}
bar_fig_biome$percent_biome_label <-percent(bar_fig_biome$percent_biome_sm)

#reorder sytems
bar_fig_biome$system_sm <- as.factor(bar_fig_biome$system_sm)
levels(bar_fig_biome$system_sm)
bar_fig_biome$system_sm <- factor(bar_fig_biome$system_sm, 
                                  levels = c("Area under community-based natural resources management",
                                             "Area under payment for ecosystem services program",
                                             "Buffer zone",
                                             "Eco-certified production area",
                                             "Indigenous territory",
                                             "IPLC-led conservation area",
                                             "Military training ground",
                                             "State-designated protected area",
                                             "State-required sustainable forest production area",
                                             "Small contributors",
                                             "Other"))
levels(bar_fig_biome$system_sm) # check output

# Turn ISO into country name

bar_fig_biome$country <- factor(bar_fig_biome$country, 
                          levels = c("BOL", "BRA", "COL", "ECU", 
                                     "GUF", "GUY", "PER", "SUR", "VEN"),
                          labels = c("Bolivia", "Brazil", "Colombia", 
                                     "Ecuador", "French Guiana", "Guyana", "Peru",
                                     "Suriname", "Venezuela"))

# colors 
amazoncol_sm <- c("#8c6bb1", # "Area under community-based natural resources management"
               "#74a9cf", #"Area under payment for ecosystem services program"
               "#AEE171", #"Buffer zone"
               "#CD6E6E", # "Eco-certified production area"
               "#6677CD", # "Indigenous territory"
               "#6A3D9A", #"IPLC-led conservation area",
               "#CC6428", #"Military training ground"
               "#feb24c",#"small contributor"
               "#006837", #"State-designated protected area"
               "#fcc5c0", # "State-required sustainable forest production area"
               "#CCCCCC" #"Other"
)
# Amazon boundary: #686868
# boundary color: #828282 (50% gray)
# country background: white 

names(amazoncol_sm) <- c(
  "Area under community-based natural resources management", 
  "Area under payment for ecosystem services program",
  "Buffer zone",
  "Eco-certified production area",
  "Indigenous territory",
  "IPLC-led conservation area",
  "Military training ground",
  "Small contributors",
  "State-designated protected area",
  "State-required sustainable forest production area",
  "Other") 

ggplot(bar_fig_biome, aes(x = system_sm, y = percent_biome_sm, fill = system_sm))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = amazoncol_sm)+
  facet_wrap(~country, nrow = 3, ncol = 3)+
  geom_text(aes(label = percent_biome_label), 
            position = position_dodge(0.9),
            vjust = - 0.5,
            size = 2.5)+
  ylab("Percent of country's total Amazon biome area")+
  guides(fill=guide_legend(title="System subfamily"))+ # change legend title 
  scale_y_continuous(labels = scales::percent)+
  theme(axis.title.x=element_blank(), # remove x labels - too long 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())



```

#Figure S3.  Total area, in km2, and percentage of country land area under area-based conservation governance system by country, excluding data represented as points and statistics.

```{r}


land_ctr <- rename(land_ctr, "country" = "ISO_CODE")
land_ctr <- land_ctr %>% 
  select(country, area_geo)

names(land_by_ctr)[2] <- "country"

ctr_land_bar <- land_ctr %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>% 
  right_join(land_by_ctr, by = "country") 

# calculate percent land
ctr_land_bar$percent_land <- ctr_land_bar$area_geo/ctr_land_bar$land_area 

ctr_land_bar$percent_land_label <-paste("(", percent(ctr_land_bar$percent_land), ")", sep = "")

# add label for area conserved 
ctr_land_bar$area_label <-round(ctr_land_bar$area_geo, digits = 2) 

#update country name
ctr_land_bar$country <- factor(ctr_land_bar$country, 
                          levels = c("BOL", "BRA", "COL", "ECU", 
                                     "GUF", "GUY", "PER", "SUR", "VEN"),
                          labels = c("Bolivia", "Brazil", "Colombia", 
                                     "Ecuador", "French Guiana", "Guyana", "Peru",
                                     "Suriname", "Venezuela"))

ggplot(ctr_land_bar, aes(x = country, y = percent_land))+
  geom_bar(stat = "identity", fill = "steelblue")+
  geom_text(aes(label = percent_land_label), 
            position = position_dodge(0.9),
            vjust = - 0.5,
            size = 2.5)+
  geom_text(aes(label = area_label), 
            position = position_dodge(0.9),
            vjust = - 2.5,
            size = 2.5)+
  annotate("text", x = 8, y = 0.85, label = 
  "Total land under area-based 
   conservation governance system: 
   5,566,753 km2 (40.96%) ", size = 3)+
  ylab("Percent of country's total land area")+
  scale_y_continuous(labels = scales::percent)


# sum(land_ctr$area_geo) / totallandv
# percent land conserved: 5566753/13591776
```

#Figure S4. Total area, in km2, and percentage of Amazon biome under area-based conservation governance systems by country.

```{r}

biome_by_ctr  # a country's amazon biome area

biome_ctr <- rename(biome_ctr, "country" = "ISO_CODE") # a country's amazon biome under conservation system

ctr_biome_bar <- biome_ctr %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>% 
  right_join(biome_by_ctr, by = "country")

# calculate percent biome
ctr_biome_bar$percent_biome <- ctr_biome_bar$area_geo/ctr_biome_bar$biome_area 
ctr_biome_bar$percent_biome_label <-paste("(", percent(ctr_biome_bar$percent_biome), ")", sep = "")

# add label for area conserved 
ctr_biome_bar$area_label <-round(ctr_biome_bar$area_geo, digits = 2) 

#update country name
ctr_biome_bar$country <- factor(ctr_biome_bar$country, 
                          levels = c("BOL", "BRA", "COL", "ECU", 
                                     "GUF", "GUY", "PER", "SUR", "VEN"),
                          labels = c("Bolivia", "Brazil", "Colombia", 
                                     "Ecuador", "French Guiana", "Guyana", "Peru",
                                     "Suriname", "Venezuela"))

ggplot(ctr_biome_bar, aes(x = country, y = percent_biome))+
  geom_bar(stat = "identity", fill = "steelblue")+
  geom_text(aes(label = percent_biome_label), 
            position = position_dodge(0.9),
            vjust = - 0.5,
            size = 2.5)+
  geom_text(aes(label = area_label), 
            position = position_dodge(0.9),
            vjust = - 2.5,
            size = 2.5)+
  annotate("text", x = 8, y = 0.85, label = 
  "Total Amazon biome under 
   area-based conservation 
   governance system: 
   4,340,711.69 km2 (62.44%)", size = 3)+
  ylab("Percent of country's total Amazon biome area")+
  scale_y_continuous(labels = scales::percent)

# percent biome conserved = 4340711.693/6951904.10 (using biome area after removing no country areas )
```


# Figure S5. Percent of country’s total land area that belongs to each subfamily, by country. 

```{r }

# vertical
ggplot(bar_fig, aes(system_sm, percent_land_sm, fill = country))+
  geom_bar(position = "dodge",
           stat = "identity")+
  scale_fill_brewer(palette = "Set3")+
  coord_flip()+
  ylab("Percent of country's total land area")+ 
  scale_y_continuous(labels = scales::percent)+
  aes(stringr::str_wrap(system_sm, 15))+
  xlab(NULL)+
  geom_text(aes(label = percent_land_label), 
            position = position_dodge(0.9),
            hjust = - 0.3,
            size = 2.5)

```


# Table S3. Total area in km2 of both terrestrial and marine area-based conservation governance system by country, excluding data represented as points and statistics. 

```{r}
#ctr_sys
names(ctr_sys)[2] <- "Country"
names(ctr_sys)[1] <- "System" 
ctr_sys_tbl <- ctr_sys %>%
  clean_names() %>%
  select(country, system, area_geo) %>% 
  filter(country %in% c('BOL', 'BRA', 'COL', 'ECU', 'GUF', 'GUY', 'PER', 'SUR', 'VEN'))%>% 
  mutate(across(where(is.numeric), round, digits = 2)) %>% 
  pivot_wider(names_from = country, values_from = area_geo) %>% 
  select(system, BOL, BRA, COL, ECU, GUF, GUY, PER, SUR, VEN) # reorder columns

# add row sum
ctr_sys_tbl <- ctr_sys_tbl %>% 
  mutate(sum = rowSums(.[,2:10], na.rm = T))

#create 2 list used repeatedly in later commands
levelist <- c("Area under community-based natural resources management",
              "Area under payment for ecosystem services program", 
              "Buffer zone", 
              "Conservation agreement",
              "Conservation concession", 
              "Eco-certified production area", 
              "Indigenous territory", 
              "IPLC-led conservation area", 
              "Military training ground", 
              "Private recreational area", 
              "Privately protected area",
              "State-designated protected area", 
              "State-designated recreational area",
              "State-required sustainable forest production area", 
              "Other")
colnamelist <- c("System subfamily", "Bolivia", "Brazil", "Colombia", 
                        "Ecuador", "French Guiana", "Guyana", "Peru",
                        "Suriname", "Venezuela", "Total")

# reorder rows
ctr_sys_tbl$system <- as.factor(ctr_sys_tbl$system)
levels(ctr_sys_tbl$system)
ctr_sys_tbl$system <- factor(ctr_sys_tbl$system,
                             levels = levelist)
levels(ctr_sys_tbl$system) # check output

ctr_sys_tbl<- with(ctr_sys_tbl, ctr_sys_tbl[order(system),])

# rename columns 
names(ctr_sys_tbl) <- colnamelist

t1 <- ctr_sys_tbl %>% 
  flextable()
t1 <- theme_vanilla(t1)
t1 <- add_footer_lines(t1, "Spatial overlap exists between systems.")
t1 <- color(t1, part = "footer", color = "#666666")
t1

```


# Table S4. Total area in km2 of terrestrial area-based conservation governance system by country, excluding data represented as points and statistics.

```{r }

# total system area and its percentage of land (for the table, added manually)
land_sys <- rename(land_sys, system = SYS_CAT_NEW, sys_area = area_geo)
totallandv <- as.numeric(totalland[1,2])
land_sys$percent_land <- land_sys$sys_area/totallandv
land_sys$sys_area <- round(land_sys$sys_area, digits = 2)
percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}
land_sys$percent_land_label <-
  percent(land_sys$percent_land)
land_sys$output_total <- paste(
  land_sys$sys_area, 
  "(", 
  land_sys$percent_land_label, 
  ")")
land_sys

# calculate percent land

percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}
ctr_sys_land_bar$percent_land_label <-
  percent(ctr_sys_land_bar$percent_land)

# removed non-BPA
ctr_sys_land_bar <- 
  ctr_sys_land_bar[!ctr_sys_land_bar$system == "non-BPA",]

# new column that includes area and %
ctr_sys_land_bar$results <- paste(
  ctr_sys_land_bar$area_geo, 
  "(", 
  ctr_sys_land_bar$percent_land_label, 
  ")")

ctr_sys_tbl <- ctr_sys_land_bar %>% 
  select(country, system, results) %>% 
  pivot_wider(names_from = country, values_from = results) %>% 
  select(system, BOL, BRA, COL, ECU, GUF, GUY, PER, SUR, VEN)

# reorder rows
ctr_sys_tbl$system <- as.factor(ctr_sys_tbl$system)
levels(ctr_sys_tbl$system)
ctr_sys_tbl$system <- factor(ctr_sys_tbl$system, 
                                  levels = levelist)
levels(ctr_sys_tbl$system) # check output

ctr_sys_tbl<- with(ctr_sys_tbl, ctr_sys_tbl[order(system),])

# add total system area
land_sys <-land_sys %>% 
  select(system, output_total)
ctr_sys_tbl <- left_join(ctr_sys_tbl, land_sys, by = "system" )

# rename columns 

names(ctr_sys_tbl) <- colnamelist



t2 <- ctr_sys_tbl %>% 
  flextable()
t2 <- theme_vanilla(t2)
t2 <- add_footer_lines(t2, "Spatial overlap exists between systems.")
t2 <- color(t2, part = "footer", color = "#666666")
t2


```

# Table S4 addition: land area
manually add to table S4
```{r }
land_by_ctr$country <- c("BOL", "BRA", "COL", "ECU", "GUF", "GUY", "PER", "SUR", "VEN")
land_by_ctr2 <-land_by_ctr %>% 
  select(country, land_area) %>% 
  pivot_wider(names_from = country, values_from = land_area)
t <- land_by_ctr2 %>% 
  flextable()
t <- theme_vanilla(t)
t
```


# Table S5. Total area in km2 of area-based conservation governance systems within the Amazon biome by country, excluding data represented as points and statistics. 

```{r}

# total system area and its percentage of biome
biome_sys <- rename(biome_sys, "system" = "SYS_CAT_NEW", "sys_area" = "area_geo")
total_biome_area <- sum(biome_by_ctr$biome_area) # use sum of country biome area instead of direclty from biome shp, because biome shp countains water/ocean in the amazon that is not counted in country area

biome_sys$sys_area <- round(biome_sys$sys_area, digits = 2)

biome_sys$percent_biome <- biome_sys$sys_area/total_biome_area 
percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}
biome_sys$percent_biome_label <-
  percent(biome_sys$percent_biome)
biome_sys$total_output <- paste(
  biome_sys$sys_area, 
  "(", 
  biome_sys$percent_biome_label, 
  ")")


# calculate percent biome

ctr_sys_biome_bar$percent_biome_label <-percent(ctr_sys_biome_bar$percent_biome)

# removed non-BPA.  
ctr_sys_biome_bar <- ctr_sys_biome_bar[!ctr_sys_biome_bar$system == "non-BPA",] 


# new column that includes area and %
ctr_sys_biome_bar$results <- paste(
  ctr_sys_biome_bar$area_geo, 
  "(", 
  ctr_sys_biome_bar$percent_biome_label, 
  ")")

# reorder columns
ctr_sys_biome_table <- ctr_sys_biome_bar %>%
  select(country, system, results) %>% 
  pivot_wider(names_from = country, values_from = results) %>% 
  select(system, BOL, BRA, COL, ECU, GUF, GUY, PER, SUR, VEN)

# reorder rows
ctr_sys_biome_table$system <- as.factor(ctr_sys_biome_table$system)
levels(ctr_sys_biome_table$system)
ctr_sys_biome_table$system <- factor(ctr_sys_biome_table$system, 
                                  levels = levelist)
levels(ctr_sys_biome_table$system) # check output

ctr_sys_biome_table<- with(ctr_sys_biome_table, ctr_sys_biome_table[order(system),])

# join total percent biome
biome_sys <- biome_sys %>% 
  select(system, total_output)

ctr_sys_biome_table <- left_join(ctr_sys_biome_table, biome_sys, by = "system")
# rename columns 
names(ctr_sys_biome_table) <- colnamelist

t <- ctr_sys_biome_table %>% 
  flextable()
t <- theme_vanilla(t)
t <- add_footer_lines(t, "Spatial overlap exists between systems.")
t <- color(t, part = "footer", color = "#666666")
t

```

## Table S5 addition: biome area
 manually add to table  S5
```{r}
biome_by_ctr$country <- c("BOL", "BRA", "COL", "ECU", "GUF", "GUY", "PER", "SUR", "VEN")
biome_by_ctr2 <-biome_by_ctr %>% 
  select(country, biome_area) %>% 
  pivot_wider(names_from = country, values_from = biome_area)
t <- biome_by_ctr2 %>% 
  flextable()
t <- theme_vanilla(t)
t
```



# Table S6. Area of spatial overlap between subfamilies, in km2, excluding data represented as points and statistics. 

```{r}

# keep 2 digits
overlap_area <- overlap_area %>% 
    mutate(across(where(is.numeric), round, digits = 2))  # convert to 2 digits
t1 <- overlap_area %>% 
  flextable()
t1 <- theme_vanilla(t1)
t1
```

#Table S8. Statistics and qualitative data that are not captured in Main Figure 1, Suppl. Figure 1-4, or Suppl. Table 1-5, as spatial data were not available. 
```{r}

t <- non_poly %>% 
  flextable()
t <- theme_vanilla(t)
t <- add_footer_lines(t, "Spatial overlap may exist between systems.")
t <- color(t, part = "footer", color = "#666666")
t
```

# Table S9.  Comparison of data covered in our inventory in 2016 and data reported to WDPA and WD-OECM in September 2023.

```{r}
wdpa_wdoecm_2023_matrix<- wdpa_wdoecm_2023_matrix %>% 
  mutate_all(as.character)

wdpa_wdoecm_2023_matrix <- wdpa_wdoecm_2023_matrix %>% 
  mutate_all(~ case_when(
    .== "10" ~ "Polygon",
    .== "11"~ "Polygon and point",
    .== "1" ~ "Point",
    .default = as.character(.)))


colorder <- c("ISO3",
              "Buffer zone", 
              "Conservation concession", 
              "Indigenous territory", 
              "IPLC-led conservation area", 
              "Privately protected area",
              "State-designated protected area", 
              "State-designated recreational area",
              "State-required sustainable forest production area", 
              "International designation",
              "Other",
              "non-BPA",
              "potential BPA not included in Atlas",
              "Proposed PA")
wdpa_wdoecm_2023_matrix <- wdpa_wdoecm_2023_matrix %>% 
  select(all_of(colorder)) %>% 
  arrange(ISO3)

# transpose
wdpa_wdoecm_2023_matrix_transpose <- t(wdpa_wdoecm_2023_matrix)
wdpa_wdoecm_2023_matrix_transpose <- wdpa_wdoecm_2023_matrix_transpose %>% 
  as.data.frame(row.names(wdpa_wdoecm_2023_matrix_transpose))
  
t3 <- wdpa_wdoecm_2023_matrix %>% 
  flextable()
t3 <- theme_vanilla(t3)
t3
#write.csv(wdpa_wdoecm_2023_matrix_transpose, "data/wdpa_wdoecm_2023_matrix_cleaned.csv", row.names = T)

# mannually transpose and combine with WDPA matrix in Excel. used those instead of t3 in the manuscript. 

```




## calculate small contributor percentage

```{r}
biomeArea <- sum(biome_by_ctr$biome_area)  #excluding area that did not overlap with country polygons
landArea <- as.numeric(totalland[1,2])
totalConserved <- as.numeric(total_conserved[1,2])  

sys_biome_smallcontri <- ctr_sys_biome %>% group_by(System) %>% 
  summarise(sys_area = sum(area_geo))
sys_biome_smallcontri$percentBiome <- round(sys_biome_smallcontri$sys_area/biomeArea, digits = 3)
View(sys_biome_smallcontri)


sys_land_smallcontri <- ctr_sys_land %>% 
  group_by(System) %>% 
  summarise(sys_area = sum(area_geo))
sys_land_smallcontri$percentLand <- round(sys_land_smallcontri$sys_area/landArea, digits = 3)
View(sys_land_smallcontri)

sys_totalarea_smallcontri <- ctr_sys %>% 
  group_by(System) %>% 
  summarise(sys_area = sum(area_geo))
sys_totalarea_smallcontri <- sys_totalarea_smallcontri %>% 
  filter(System%in% c("Private recreational area", 
                            "Conservation concession", 
                            "Conservation agreement", 
                            "Privately protected area", 
                            "State-designated recreational area"))
sys_totalarea_smallcontri_percent <- sum(sys_totalarea_smallcontri$sys_area)/totalConserved
sys_totalarea_smallcontri_percent


```
Collectively contribute to <1% of total area conserved: 
Private recreational area; Conservation concession; Conservation agreement; Privately protected area; State-designated recreational area 

